Refinement for open automata

### 2021-05-31 2021-Refinement/report.tex
Added comments in source to not pretend the proof is complete

For three open automata \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\), such that \(R_{12}\) is a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\), we want to define \(R_{13}\) by saying that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier as we did in deadlock reduction, which is possible because states are finite.
\nmm{R_{13} \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}} is a witness of the transitivity, proof is given in appendix \ref{apx:trans}.
\begin{thm}[Context refinement] Hole-equal simulation is context refining \end{thm}
\begin{proof}
% TODO: assume a <= b & a[c] -> b[c] + proof
\end{proof}
% Rédiger un paragraphe sur le fait que composer dans l'implem sans deadlock implique qu'il n'y a pas de deadlock quand on compose avec la spec, on voudrais avoirs l'autre sens mais ce n'est pas possible
\begin{prop}[Composition correctness'] Hole-subset simulation is correct w.r.t.\@ composition. \end{prop}
\begin{proof}
% TODO
\end{proof}
\begin{prop}[Composition correctness"]
Let \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3\) be open automata.
If \(R\) is a hole-superset simulation of \(A_1\) by \(A_2\) then let \(j \in J_1 \setminus J_2\), there is a hole-superset simulation of \(A_1\subst{A_3}\) by \(A_2\).
\end{prop}
\begin{proof}
% TODO
\end{proof}
% J'2 \subseteq J'1 in hole-subset, J'1 \subseteq J'2 in hole-superset (decide of transition variable with holes); composition with new holes is hole-subset where hole variables are free then hole-superset where hole names have no conflicts with removed holes. Transitivity must ensure holes name have no conflict for that reason!

@@

\defitem We prove that initial configurations are related using \(\sigma_{02}\) as the values of the variables in \(V_2\), and \(s_2\) as the branch of the disjunction.

@@

	\item[\goal{0}:] Let \nmm{t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}}.

@@

	% TODO HERE: goal 3

@@

@@

		V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R_{12}\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \hyp{6} \\
		V_2 \uplus V_3 \uplus \biguplus_\subbox{t_3 \in \fOT{s_3}} \fvars{t_3} \vdash R_{23}\mpar{s_2, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \implies \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \hyp{7} \\

@@

	\[ \sigma_1 \uplus \sigma_3 \vdash R_{13}\mpar{s_1, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \hyp{5} \]

@@

		\sigma_1 \uplus \sigma_2 \uplus \sigma_3 & \vdash R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \hyp{5'} \\
		&& \sigma_1 \uplus \sigma_3 & \vdash \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \hyp{5"}

@@

	We can use immediately \hyp{7} with valuation \(\sigma_2 \uplus \sigma_3\) to get \(\nu_2: \biguplus_{t_2 \in \fOT{s_2}} \fvars{t_2} \to \values\).
	The left side of the implication is proved with the right side of \hyp{5'} and \hyp{5"}, so we get
	\[ \sigma_2 \uplus \sigma_3 \uplus \nu_2 \vdash \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \hyp{7'} \]
	We can now use \hyp{6} with valuation \(\sigma_1 \uplus \sigma_2 \uplus \nu_2\) to get \(\nu_1: \biguplus_{t_1 \in \fOT{s_1}} \fvars{t_1} \to \values\).
	The left side of the implication is proved with the left side of \hyp{5'} and \hyp{7'}, so we get
	\[ \sigma_1 \uplus \sigma_2 \uplus \nu_1 \vdash \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \hyp{6'} \]
	Finally we can prove \goal{1'} with valuation \(\nu_1\) fot the implicit \(\exists\) by admitting the left side of the implication and using \hyp{6'} for the right side.

>>>>

For three open automata \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\), such that \(R_{12}\) is a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\), we want to define \(R_{13}\) by saying that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier as we did in deadlock reduction, which is possible because states are finite.
\nmm{R_{13} \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}} is a witness of the transitivity, proof is given in appendix \ref{apx:trans}.
\begin{thm}[Context refinement] Hole-equal simulation is context refining \end{thm}
\begin{proof}
% TODO: assume a <= b & a[c] -> b[c] + proof
\end{proof}
% Rédiger un paragraphe sur le fait que composer dans l'implem sans deadlock implique qu'il n'y a pas de deadlock quand on compose avec la spec, on voudrais avoirs l'autre sens mais ce n'est pas possible
\begin{prop}[Composition correctness'] Hole-subset simulation is correct w.r.t.\@ composition. \end{prop}
\begin{proof}
% TODO
\end{proof}
\begin{prop}[Composition correctness"]
Let \(A_1 \defobject \OAg[1]\) be hole-superset simulated by \(A_2 \defobject \OAg[2]\).
Let \(j \in J_1 \setminus J_2\) and \(A_3\) an open automaton.
There is a hole-superset simulation of \(A_1\subst{A_3}{j}\) by \(A_2\).
\end{prop}
\begin{proof}
% TODO
\end{proof}
% J'2 \subseteq J'1 in hole-subset, J'1 \subseteq J'2 in hole-superset (decide of transition variable with holes); composition with new holes is hole-subset where hole variables are free then hole-superset where hole names have no conflicts with removed holes. Transitivity must ensure holes name have no conflict for that reason!

@@

\defitem We prove that initial configurations are related using \(\sigma_{02}\) as the values of the variables in \(V_2\), and \(s_{02}\) as the branch of the disjunction.

@@

	\item[\goal{0}:] Let \nmm{t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}}. % TODO: quantify on s_2

@@

	Let \(\sigma_1: V_1 \uplus \fvars{t_1} \to \values\) and \(\sigma_3: V_3 \to \values\).
	We use dummy values for variables in \nmm{\biguplus_\subbox{\mpar{x, y} \in Z} \fvars{t_{3xy}}} to get the left side of this implication as an hypothesis (independant from these values):
	\[ \sigma_1 \uplus \sigma_3 \vdash R_{13}\mpar{s_1, s_3} \wedge g_1 \hyp{2} \]
		\sigma_1 \uplus \sigma_2 \uplus \sigma_3 & \vdash R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \hyp{2'} \\
		&& \sigma_1 & \vdash g_1 \hyp{2"}
	We can use immediately \hyp{0"} with valuation \(\sigma_1 \uplus \sigma_2\) to get \nmm{\nu_2: \biguplus_{x \in X} \fvars{t_{2x}} \to \values}.
	The left side of the implication is proved with the right side of \hyp{2'} and \hyp{2"}, so we get \(x \in X\) and the hypotheses
		\sigma_1 \uplus \sigma_2 \uplus \nu_2 & \vdash \alpha_1 = \alpha_{2x} \hyp{3a} \\
		\sigma_1 \uplus \sigma_2 \uplus \nu_2 & \vdash \bigwedge_\subbox{j \in J'_1 \cap J'_{2x}} \beta_{1j} = \beta_{2xj} \hyp{3b} \\
		\sigma_2 \uplus \nu_2 & \vdash g_{2x} \hyp{3c} \\
		\sigma_1 \uplus \sigma_2 \uplus \nu_2 & \vdash R\mpar{s'_1, s'_{2x}}\psubst{\psi_1 \uplus \psi_{2x}} \hyp{3d}
	We can now use \hyp{1x'} with valuation \(\sigma_2 \uplus \nu_2 \uplus \sigma_3\) to get \nmm{\nu_3: \biguplus_{y \in Y_x} \fvars{t_{3xy}} \to \values}.
	The left side of the implication is proved with the left side of \hyp{2'} and \hyp{2"}, so we get \(y \in Y_x\) and the hypotheses
		\sigma_2 \uplus \nu_2 \uplus \sigma_3 \uplus \nu_3 & \vdash \alpha_{2x} = \alpha_{3xy} \hyp{4a} \\
		\sigma_2 \uplus \nu_2 \uplus \sigma_3 \uplus \nu_3 & \vdash \bigwedge_\subbox{j \in J'_{2x} \cap J'_{3xy}} \beta_{2xj} = \beta_{3xyj} \hyp{4b} \\
		\sigma_3 \uplus \nu_3 & \vdash g_{3xy} \hyp{4c} \\
		\sigma_2 \uplus \nu_2 \uplus \sigma_3 \uplus \nu_3 & \vdash R\mpar{s'_{2x}, s'_{3xy}}\psubst{\psi_{2x} \uplus \psi_{3xy}} \hyp{4d}
	Finally to prove \goal{3} with valuation \(\nu_3\) for the implicit \(\exists\), we admit the left side of the implication choose to prove the branch \(\mpar{x, y}\):
		\sigma_1 \uplus \sigma_3 \uplus \nu_3 & \vdash \alpha_1 = \alpha_{3xy} \goal{3a} \\
		\sigma_1 \uplus \sigma_3 \uplus \nu_3 & \vdash \bigwedge_\subbox{j \in J'_{3xy} \cap J_1} \beta_{1j} = \beta_{3xyj} \goal{3b} \\
		\sigma_3 \uplus \nu_3 & \vdash g_{3xy} \goal{3c} \\
		\sigma_1 \uplus \sigma_3 \uplus \nu_3 &\vdash R_{13}\mpar{s'_1, s'_{3xy}}\psubst{\psi_1 \uplus \psi_{3xy}} \goal{3d}
	\goal{3a} is proved by transitivity using \hyp{3a} and \hyp{4a}.
	\goal{3b} depends on \(\triangle\).
	If \(\triangle\) is \(\subseteq\) % TODO
	If \(\triangle\) is \(\supseteq\) % TODO
	\goal{3c} is exactly \hyp{4c}.
	\goal{3d} % TODO: branch s_2

@@

	We have the hypotheses on \(R_{12}\) and \(R_{23}\), then the current goal: % TODO HERE: get s_2 by decomposing R_{13}

@@

		V_1 \uplus V_2 \uplus \biguplus_\subbox{t_2 \in \fOT{s_2}} \fvars{t_2} \vdash R_{12}\mpar{s_1, s_2} \wedge \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \implies \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \hyp{5} \\
		V_2 \uplus V_3 \uplus \biguplus_\subbox{t_3 \in \fOT{s_3}} \fvars{t_3} \vdash R_{23}\mpar{s_2, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \implies \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \hyp{6} \\

@@

	\[ \sigma_1 \uplus \sigma_3 \vdash R_{13}\mpar{s_1, s_3} \wedge \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \hyp{7} \]

@@

		\sigma_1 \uplus \sigma_2 \uplus \sigma_3 & \vdash R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3} \hyp{7'} \\
		&& \sigma_1 \uplus \sigma_3 & \vdash \bigvee_\subbox{t_3 \in \fOT{s_3}} \fguard{t_3} \hyp{7"}

@@

	We can use immediately \hyp{6} with valuation \(\sigma_2 \uplus \sigma_3\) to get \nmm{\nu_2: \biguplus_{t_2 \in \fOT{s_2}} \fvars{t_2} \to \values}.
	The left side of the implication is proved with the right side of \hyp{7'} and \hyp{7"}, so we get
	\[ \sigma_3 \uplus \nu_2 \vdash \bigvee_\subbox{t_2 \in \fOT{s_2}} \fguard{t_2} \hyp{6'} \]
	We can now use \hyp{5} with valuation \(\sigma_1 \uplus \sigma_2 \uplus \nu_2\) to get \nmm{\nu_1: \biguplus_{t_1 \in \fOT{s_1}} \fvars{t_1} \to \values}.
	The left side of the implication is proved with the left side of \hyp{7'} and \hyp{6'}, so we get
	\[ \sigma_1 \uplus \nu_1 \vdash \bigvee_\subbox{t_1 \in \fOT{s_1}} \fguard{t_1} \hyp{5'} \]
	Finally we can prove \goal{1'} with valuation \(\nu_1\) for the implicit \(\exists\) by admitting the left side of the implication and using \hyp{5'} for the right side.
